---
title: 'Tidy Tuesday: San Francisco Tree'
output: 
  html_document:
    code_folding: "hide"
    fig_width: 9
    fig_height: 5
---

```{r global options, include = FALSE}


library(tidytuesdayR)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(mapdeck)


knitr::opts_chunk$set(comments = FALSE, warning=FALSE, message=FALSE)
```

### General Set Up / Process

- Set custom `ggplot2` theme
- Set custom color scale
- Load data using `tidytuesdayR` package
- _Slightly_ modify the dataset
  - Calculate approximate age of tree
  - Replace odd (>200) tree diameters with average diameter
  - Filter out trees with lat and long not _actually_ in San Francisco 
  

``` {r}

# Set Themes --------------------------------------------------------------
theme_set(theme_classic())

theme_minimal() %+replace% 
  theme(
  axis.title.y = element_blank(), 
  axis.title.x = element_blank(),
  panel.grid.minor.x = element_blank(),
  panel.grid.major.x = element_blank(),
  legend.title = element_blank(),
  complete = T
) %>%
  theme_set()

colors <-  colorRampPalette(c("#A7B59F", "#153A00"))(6)

# Load Data Sets ----------------------------------------------------------
tt_data <- tidytuesdayR::tt_load('2020-01-28') 
sf_trees <- tt_data$sf_trees


# A little bit of clean up ------------------------------------------------
# Calculate appro AGE in years
# Remove unusual diameter (dbh) values

sf_trees <- sf_trees %>%
  mutate(
    age = interval(date, today()) / years(1),
    dbh = ifelse(dbh > 200 | dbh == 0, NA, dbh)
  )
```



### Explore Diameter and Age Measures

```{r}
plot_dist_diameter <- ggplot(sf_trees, aes(dbh)) +
  geom_histogram(binwidth = 5, fill = colors[6], alpha = .75, color = '#FFFFFF') +
  scale_y_continuous(labels = scales::comma) +
  labs(title = 'SF Trees Diameter Distribution')

plot_dist_diameter
```

```{r}
plot_dist_age <- ggplot(sf_trees, aes(age))+
  geom_histogram(binwidth = 5, fill = colors[6], alpha = .75, color = '#FFFFFF') +
  scale_y_continuous(labels = scales::comma) +
  labs(title = 'SF Trees Age Distribution')

plot_dist_age
```

```{r}
plot_diam_by_age <- ggplot(sf_trees, aes(date,dbh)) +
  geom_hex(alpha = .95, color = '#FFFFFF') +
  scale_fill_gradient(low='#E2E6DF', high = colors[6]) +
  labs(title = 'SF Trees Diameter by Date')

plot_diam_by_age
```

### Create Functions for Mapping

- Filter unusable lat and long 
- Summarize stats
  - Attempt to summary stats into 100 square meter areas
  - Uncount dataframe as a workaround since `Mapdeck` scales and colors based on bin counts

```{r}
# Create Prep and Map Functions -------------------------------------------

f_prepare_data <- function(df, measure) {
  
  ## Remove lat and long not close to SF
  df <- df %>%
    filter(
      !is.na(latitude), 
      !is.na(longitude),
      latitude >= 37.7,
      latitude < 38,
      longitude >= -123,
      longitude <= -122
    ) 
  
  ### Round Lat and Long to 3rd decimal and summarize 
  df <- df %>%
    select(latitude, longitude, measure = !!measure) %>%
    filter(!is.na(measure)) %>%
    mutate(
      latitude = round(latitude,3),
      longitude = round(longitude,3)
    ) %>%
    group_by(latitude, longitude) %>%
    summarize(mean = mean(measure, rm.na = T)) %>%
    mutate(mean = replace_na(mean, 0)) 
  
  ## Uncount mean. 
  ## Necessary transform in order to map appropriately
  df %>%
    uncount(mean)
  
}

plot_map <- function(df) {

  df %>%
    mapdeck(
      token = Sys.getenv('MAPBOX_TOKEN'),
      style = mapdeck_style("light"),
      pitch = 55,
      zoom = 12,
      bearing = 10,
      location = c(-122.445,37.75)
    ) %>%
      add_hexagon(
        lat = "latitude",
        lon = "longitude",
        layer_id = "hex_layer",
        elevation_scale = 2,
        radius = 11.132^2,
        colour_range = colors,
        highlight_colour = '#E2E6DFFF',
        auto_highlight = T,
        update_view = F
      ) 
  
}
```

### Map: Average Tree Diameter

Taller and darker bins equate to larger trees on average.

```{r}
sf_trees %>%
  f_prepare_data(quo(dbh)) %>%
  plot_map()
```

### Map: Average Tree Age

Taller and darker bins equate to older trees on average.

```{r}
sf_trees %>%
  f_prepare_data(quo(age)) %>%
  plot_map()
```

### Notes

Libraries

- bootstraplib
- tidytuesdayR
- tidyverse
- ggplot2
- lubridate
- mapdeck

Dataset: [TidyTuesday Data](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-01-28)